Gyrochronal and Lithium Ages for Stars and Planets Observed by Kepler

### 2023-06-07 paper/ms.tex
update ms

\begin{abstract}
  We report rotation- and lithium-based ages for stars and planets observed by
  the primary Kepler mission.  Stellar rotation periods are collected from
  previous analyses of the Kepler light curves; lithium content is assessed
  using HIRES spectra.
  We report ages for \nkoisgoldwithages\ FGK stars with transiting planets, with relative statistical
  precisions $\sigma_t / t$ of XX +YY -ZZ\%; a factor of X.X
  +Y.Y -Z.Z more precise than state-of-the-art isochronal ages for the
  same stars.  The results include XXX planets younger than 1\,Gyr,
  and YYY planets younger than 0.5\,Gyr.
  In the base sample of dwarf FGK stars observed by
  Kepler, we find that the intrinsic age distribution is(/not)
  consistent with a uniform distribution between X and Y billion years
  (Gyr).  In the planet sample, we find a consistent(/inconsistent)
  distribution at $N$-sigma.  Analyzing the completeness of the Kepler
  detection pipeline, we interpret the majority of the change to be
  caused by selection effects (or astrophysics).  We also comment on a
  few previously reported evolutionary trends in the size evolution of
  close-in mini-Neptunes (2-4\,\rearth) and super-Earths
  (1-2\,\rearth), a few of which we recover at comparable
  ($\sim$3$\sigma$) significance to previous work, and one of which
  we note for the first time.
\end{abstract}

>>>>

% <250 words
\begin{abstract}
  We report rotation- and lithium-based ages for stars and planets observed by
  the primary Kepler mission.  Stellar rotation periods are collected from
  previous analyses of the Kepler light curves; lithium content is assessed
  using HIRES spectra.
  We report ages for \nkoisgoldwithages\ FGK stars with transiting
  planets, with relative statistical precisions $\sigma_t / t$ of XX
  +YY -ZZ\%; a factor of X.X +Y.Y -Z.Z more precise than
  state-of-the-art isochronal ages for the same stars.
  The results include XXX planets younger than 1\,Gyr (at 2$\sigma$),
  and YYY planets younger than 0.5\,Gyr (at 2$\sigma$).
  In the base sample of dwarf FGK stars observed by Kepler, we find
  that the intrinsic age distribution is inconsistent with a uniform
  distribution, and instead rises roughly linearly from 0-3\,Gyr;
  stars 2\,Gyr old are twice are common in the Kepler field as stars
  1\,Gyr old.
  This agrees with results from previous isochronal analyses, and extends
  beyond their age sensitivity, down to $\approx$0.1\,Gyr.
  %TODO check A-D test
  The sample of Kepler planets has a (in)consistent age distribution.
  Analyzing the completeness of the Kepler
  detection pipeline, we interpret the majority of the change to be
  caused by selection effects (or astrophysics).  We also comment on a
  few previously reported evolutionary trends in the size evolution of
  close-in mini-Neptunes (2-4\,\rearth) and super-Earths
  (1-2\,\rearth), a few of which we recover at comparable
  ($\sim$3$\sigma$) significance to previous work, and one of which
  we note for the first time.
\end{abstract}
